Higher-Order unification for free!

### 2024-04-16 paper/main.tex
Update proof th 5.18 and th 6.4

\begin{theorem}[Fidelity in $\wellb_\beta$] \Fullref{prop:simulation} and
\fullref{prop:fidelity} hold if $\;\wellb_\beta(\foUnifPb)$
\end{theorem}
\begin{proof}[Proof sketch] thanks to \cref{lem:w-enforcement-maybebeta}
  \Ue is as powerful as \Uo in \maybebeta, as well as in \wellb by
  \cref{prop:uniffowb}.
\end{proof}

@@

\begin{theorem}[Fidelity in $\wellb_{\beta \eta}$]
  % F x = z, F x y = z y
  Given a list of unification problems \foUnifPb, if
  $\wellb_{\beta \eta}(\foUnifPb)$ and if the 
  memory map is bijective, then
  the introduction 
  of \linketa guarantees \fullref{prop:fidelity}.
  \footnote{This premise is fulfilled in \cref{sec:invariant1}}
  \label{lemma:fidelity-maybeeta}
\end{theorem}

\begin{proof}[Proof sketch]
  \progressetaleft and \progressetadedup activate a \linketa when, in the
  original unification problem, a \maybeeta term is unified with respectively a
  well-behaved term or another \maybeeta term. In both cases, the links trigger
  a unification which succeeds iff the same unification in \Fo{} succeeds,
  \todo{Fix: show that 5.7 an 5.8 compensate for missing eta rule, citare regolae in =o}
  guaranteeing \cref{prop:fidelity}. \progressetaright never fails, in fact,
  this progression refines a variable to a rigid term and plays no role in 
  \cref{prop:fidelity}.
  % If 1) a \linketa is activated by \progressetaleft, then unification is done
  % between a \maybeeta term $t_1$ and a term $t_2$ (with $\wellb(t_2)$). This
  % activation performs a unification which succeeds iff the original problem in
  % \Fo{} succeeds. If 2) a \linketa is activated by \progressetadedup, then the
  % unification is done between two \maybeeta terms, and again this unification
  % succeeds iff it succeeds in \Fo{}. Finally, if 3) a \linketa is activated by
  % \progressetaright, the unification, done between a variable and a term,
  % always succeeds, this is what we expect to guarantee fidelity, since we
  % are essentially removing a hole added by the compilation in the place of a
  % \maybeeta subterm. In all the cases fidelity is respected.
\end{proof}

@@

\begin{proof}[Proof sketch]
  By the definition of \mapdeduplication, any two occurrencies of the same \Fo{} 
  variables with different arities are compiled to two
  variables $A^n$ and $C^m$ related by a chain of \linketa.
  We show that whenever one of the two is instantiated with a rigid term then
  the chain becomes shorter and hence, eventually, the
  other variable is instantiated too.
  Without loss of generality we assume $m < n$ and $A$ is instantiated.
  The corresponding \linketa is triggered by \progressetaleft instantiating
  the next variable in the chain of links.
\end{proof}
\todo{Not sure I fullfy fixed the proof, but the old oen was not ok}

>>>>

\begin{theorem}[Fidelity in $\wellb_\beta$] \Fullref{prop:simulation} and
\fullref{prop:fidelity} hold if $\;\wellb_\beta(\foUnifPb)$.
\label{th:fidelity-beta}
\end{theorem}
\begin{proof}[Proof sketch] thanks to \cref{lem:w-enforcement-maybebeta}
  \Ue is as powerful as \Uo in \maybebeta, as well as in \wellb by
  \cref{prop:uniffowb}.
\end{proof}

@@

\begin{theorem}[Fidelity in $\wellb_{\beta \eta}$]
  % F x = z, F x y = z y
  Given a list of unification problems \foUnifPb, such that
  $\wellb_{\beta \eta}(\foUnifPb)$, if the 
  memory map is bijective, then
  the introduction 
  of \linketa guarantees \fullref{prop:fidelity}.
  \footnote{This premise is fulfilled in \cref{sec:invariant1}}
  \label{lemma:fidelity-maybeeta}
\end{theorem}

\begin{proof}[Proof sketch]
  We want to prove that for any $\fstep_i$, $\fstep_i$ succeeds iff $\hstep_i$
  succeeds. An $\hstep_i$ is made by a unification $\mathcal{U}$ and a link
  progression $\mathcal{P}$.
  Two cases should be analyzed: 1) $\fstep_i$ succeeds, this means that $\mathcal{U}$
  succeeds since any \maybeeta term is replaced with a fresh variable, and by
  \cref{th:fidelity-beta}, the other subterms unify. Moreover, $\mathcal{P}$
  succeeds, since the $\eta$-progression
  rules simply imitate the rules~\ref{clause:eta1} and \ref{clause:eta2},
  performing unification between $\eta$-contracted terms. 
  If 2) $\fstep_i$ fails, either the failure is 
  due to the unification of terms that are not in \maybeeta. In this case
  $\mathcal{U}$ fails thanks to \cref{th:fidelity-beta}. 
  On the other hand, the failure is
  due to a unification over \maybeeta terms, making $\mathcal{P}$ to fail.
\end{proof}

@@

\begin{proof}[Proof sketch]
  By the definition of \mapdeduplication, any two occurrencies of the same \Fo{} 
  variable used at different arities are compiled to two
  variables $A^m$ and $C^n$ related by a chain of \linketa.
  Whenever the two variables are assigned to the non-unifiable terms,
  the chain of \linketa is entirely waken up refining little by little
  the intermediate variables between $A^m$ and $C^n$ leading eventually to a failure. 
  On the other
  hand, if the two variables still remain unifiable after their assignement, 
  then the activation of 
  the corresponding \linketa just refines the chain, by reducing eventually
  its length.
\end{proof}
% \todo{Not sure I fullfy fixed the proof, but the old oen was not ok}
